---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "><>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fishblicc

<!-- badges: start -->
<!-- badges: end -->

The goal of fishblicc is to provide tools to fit a catch curve model to individual length frequency samples, analogous to an age-based catch curve. However, the model can account for length-based selectivity and multiple gears. The model is fitted using MCMC in Stan (mc-stan.org) to account for uncertainty. The approach could be useful for data-limited stock assessments, such as assessing stock status at end of projects where length data have been collected, or continuous monitoring of low-data bycatch species as part of a harvest strategy. The model also provides tools to assess how effective length sampling is in estimating quantities of interest.  

## Installation

You can install the development version of fishblicc like so:

``` r
if (require("remotes")) install.packages("remotes")
install_github("PaulAHMedley/fishblicc")
```

## Model Description

This is the Bayesian Length Interval Catch Curve project to estimate mortality from length frequency samples that form a snapshot of a fishery. Length samples are not treated as a time series. This is explicitly a "data-limited" method designed to be used as a risk assessment approach rather than provide definitive information on stock status. Although the model does not require a data time series, it assumes the data are a snapshot of the catch length composition from a fishery in a stationary state. It is suitable for end-of-project evaluation of length frequency data that may have been collected over a short time period (e.g. a year) or to monitor status of many species that may take up a small proportion of the catches.  
The assessment is carried out in R using Stan (mc-stan.org) to carry out the MCMC.  

In common with other such methods, there are significant assumptions that will not be met in practice, therefore the focus of this method is robustness and looking for ways to reduce the impact of model structural errors or at least spot when such errors may invalidate the results.  

The method can be used to evaluate stock status based on the spawning potential ratio. In general, this uses the ratio of large (older, mature) fish in a sample compared to small (immature young) fish. If the ratio is too small compared to what might be expected in a sample of a population that is unfished, the population will be at high risk of overexploitation.  

The model used here does not convert length to age, but instead accounts for mortality as fish grow through length intervals. The estimation method works by assuming that the von Bertalanffy growth model describes the length at age and variation in individual fish growth is governed by their maximum length which is drawn from a gamma probability density function (constant CV). The model is flexible enough to include all data, so, for example, observed lengths above $L_\infty$ are not rejected.  

The model fits the following 8 parameters:
  - Linf the asymptotic mean length  
  - Galpha the growth model inverse error parameter ($CV = Galpha^{-0.5}$)
  - Mk the natural mortality in time units of K, the growth rate  
  - Fk a scale parameter for the fishing mortality in time units of K, the growth rate  
  - Smx the maximum point for the double-sided normal selectivity model
  - Ss1 the lower-side slope parameter for the double-sided normal selectivity model (i.e. 1/variance)
  - Ss2 the upper-side slope parameter for the double-sided normal selectivity model (equal to zero for a flat-topped selectivity.
  - phi the over-dispersion parameter of the counts in the length bins for the negative binomial.  

The length catch curve is used to estimate the spawning potential ratio. To do this, estimates of natural mortality, length-weight and maturity are required. These cannot be estimated from a length frequency, so have to be provided as priors from other sources. Being data-limited, only the length-weight exponent parameter (b) and the logistic maturity at length model parameters (length at 50% maturity and logistic steepness) are required. The natural mortality is then inferred relying on life history invariants as suggested by Prince et al. Hordyk et al. (2015?). The additional parameters are:
  - $L_m$ and $L_s$ for the maturity at 50% and maturity steepness for a logistic maturity curve.  
  - $b$ parameter for the length weight relationship $W=aL^b$

## Example  

This is a basic example which shows you how to solve a common problem:  

```{r example}
library("fishblicc")
## Prepare some data in the required format
ld <- blicc_dat(LLB = 10:35,          # Lower boundaries of each length bin
          fq=c(0,1,2,5,26,70,72,66,36,34,25,24,20,
               10,12,5,3,5,6,4,2,0,2,0,1,0),    # Corresponding frequency data
          a=1.0e-4,       # Length-weight scale parameter (optional)
          b=2.95,         # Length-weight exponent
          L50=24,         # Length at 50% maturity
          Linf=c(32, 3))  # Maximum mean length normal priors
## Fit the model to these data with default settings
slim <- blicc_fit(ld)
## Calculate reference points - this takes some time
rp_df <- blicc_ref_pts(slim, ld)
## Get the model expected values by length
lx_df <- blicc_expect_len(rp_df, ld)
## Because the fit takes a little time, it might be best to save the results
save(ld, slim, rp_df, lx_df, file="fishblicc_analysis.rda") 
## Summarise the results
summary(rp_df)
plot_expected_frequency(rp_df, lx_df, ld) #Plot the results to check the model fit
```

